---
title: "Introduction to {syncdr}"
subtitle: "File Handling, Directory Comparison & Synchronization in R"
author: "Rossana Tatulli"
output: 
rmarkdown::html_vignette:
  toc: true
vignette: >
  %\VignetteIndexEntry{Introduction to {syncdr}}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

#devtools::load_all(".")
library(syncdr)
```

## Why {syncdr}?

**{syncdr}** is an R package for handling and synchronizing files and directories. Its primary objectives are:

1.  To provide a clear **snapshot of the content and status of synchronization** between two directories under comparison: including their tree structure, their common files, and files that are exclusive to either directory
2.  To make **file organization and management** in R easier: i.e., enabling content-based and modification date-based file comparisons, as well as facilitating tasks such as duplicates identification, file copying, moving, and deletion.

------------------------------------------------------------------------

ðŸ’¡\
This article does ***not*** offer a comprehensive overview of **{syncdr}** functionalities. Rather it provides a sample workflow for working with the package's main functions . After familiarizing yourself with this general workflow, read the articles throughout the rest of this website -they will explore all features of **{syncdr}** in a structured way.

------------------------------------------------------------------------

## Synchronizing with {syncdr}

**Learn how to work with {syncdr} and compare and synchronize directories in R**

Suppose you are working with two directories, let's call them `left` and `right` -each containing certain files and folders/sub-folders.

Let's first call `syncdr` function `toy_dirs()`. This generates two toy directories in `.syncdrenv` environment -say `left` and `right`- that we can use to showcase `syncdr` functionalities.

```{r toy-dirs}

# Create syncdr env with left and right directories
.syncdrenv =toy_dirs()

# Get left and right directories' paths 
left  <- .syncdrenv$left
right <- .syncdrenv$right

```

You can start by quickly comparing the two directories' tree structure by calling `display_dir_tree()`. By default, it fully recurses -i.e., shows the directory tree of all sub-directories. However, you can also specify the number of levels to recurse using the `recurse` argument.

```{r display-toy-dirs}

# Visualize left and right directories' tree structure 
display_dir_tree(path_left  = left,
                 path_right = right)

```

### Step 1: Compare Directories

The most important function in `syncdr` is `compare_directories()`. It takes the paths of left and right directories and compares them to determine their synchronization status *(see below)*. This function represents the backbone of `syncdr`: you can utilize the `syncdr_status` object it generates both:

- to *inspect* the synchronization status of files present in both directories as well as those exclusive to either directory

-   as the input for all other functions within `syncdr` that allow *synchronization* between the directories under comparison*.*

Before diving into the resulting `syncdr_status` object, note that `compare_directories()` enables to compare directories in 3 ways:

1.  By **date** only -*the default:* by default, `by_date = TRUE`, so that files in both directories are compared based on the date of last modification.

| sync_status (*all* *common files)* |
|:----------------------------------:|
| older in left, newer in right dir  |
|  newer in left, olderin right dir  |
|             same date              |

2.  By **date and content**. This is done by specifying `by_content = TRUE` (by default `by_date = TRUE` if not specifically set to FALSE). Files are first compared by date, and then only those that are newer in either directory will be compared by content.

| sync_status (*common files that are newer in either left or right, i.e., not of same date )* |
|:--------------------------------------------------------------------------------------------:|
|                                      different content                                       |
|                                         same content                                         |

3.  By **content** only, by specifying `by_date = FALSE` and `by_content = TRUE` . This option is however discouraged -comparing all files' contents can be slow and computationally expensive.

| sync_status (*all* *common files)* |
|:----------------------------------:|
|         different content          |
|            same content            |

Also, regardless of which options you choose, the sync_status of files that are exclusive to either directory is determined as:

| sync_status (*non* *common files)* |
|:----------------------------------:|
|            only in left            |
|           only in right            |

Let's now take a closer look at the output of `compare_directories()`, which is intended to contain comprehensive information on the directories under comparison. This is a list of class `syncdr_status`, containing 4 elements: (1) common files, (2) non common files, (3) left path and (4) right path

##### **1. Comparing by date**

```{r by-date}

# Compare by date only -the Default
sync_status_date <- compare_directories(left, 
                                        right)

sync_status_date
```

##### **2. Comparing by date and content**

```{r by-date-cont}

# Compare by date and content 
sync_status_date_content <- compare_directories(left, 
                                                right,
                                                by_content = TRUE)

sync_status_date_content
```

##### **3. Comparing by content only**

```{r by-content}

# Compare by date and content 
sync_status_content <- compare_directories(left, 
                                            right,
                                            by_date    = FALSE,
                                            by_content = TRUE)

sync_status_content
```

##### \*ï¸âƒ£ **Comparing directories with `verbose = TRUE`**

When calling `compare_directories()`, you have the option to enable verbose mode by setting `verbose = TRUE`. This will display both directories tree structure and, when comparing files by content, provide progress updates including the time spent hashing the files.

```{r verbose-example}

compare_directories(left,
                    right,
                    by_date    = FALSE,
                    by_content = TRUE,
                    verbose    = TRUE)

```

### Step 2: Visualize Synchronization Status

The best way to read through the output of `compare_directories()` is by visualizing it with `display_sync_status()` function.

For example, let's visualize the sync status of common files in left and right directories, when compared by date

```{r}

display_sync_status(sync_status_date$common_files,
                    left_path  = left,
                    right_path = right)
```

or let's display the sync status of non common files:

```{r}

display_sync_status(sync_status_date$non_common_files,
                    left_path  = left,
                    right_path = right)
```

### Step 3: Synchronize directories

`syncdr` enables users to perform different actions such as copying, moving, and deleting files using specific synchronization functions. Refer to the `vignette("asymmetric-synchronization")` and `vignette("symmetric-synchronization")` articles for detailed information.

For the purpose of this general demonstration, we will perform a 'full asymmetric synchronization to right'. This specific function executes the following:

-   **On common files:**
    -   If by date only (`by_date = TRUE`): Copy files that are newer in the left directory to the right directory.
    -   If by date and content (`by_date = TRUE` and `by_content = TRUE`): Copy files that are newer and different in the left directory to the right directory.
    -   If by content only (`by_content = TRUE`): Copy files that are different in the left directory to the right directory.
-   **On non common files:**
    -   Copy to the right directory those files that exist only in the left directory
    -   Delete from the right directory those files that are exclusive in the right directory (i.e., missing in the left directory)

```{r asym-sync-example}

# Compare directories

sync_status <- compare_directories(left,
                                   right,
                                   by_date = TRUE)

# Synchronize directories 
full_asym_sync_to_right(sync_status = sync_status)


```
